MSc Project - Modelling a Double Barred Galaxy as a Double Binary

### 2016-11-29 workspace/DoubleBars/MScProjectDoubleBars.tex
Started the write up

@@

where $\mathbf{r_i}=\left(x_i,y_j\right)$ and $\mathbf{r_{ij}}=\mathbf{r_j}-\mathbf{r_i}$. So for a four body problem 

@@

\ddot{\mathbf{r_1}}=\frac{m_2\mathbf{r_{12}}}{r^3_{12}}+\frac{m_3\mathbf{r_{13}}}{r^3_{13}}+\frac{m_4\mathbf{r_{14}}}{r^3_{14}}

@@

\ddot{\mathbf{r_2}}=\frac{m_1\mathbf{r_{21}}}{r^3_{21}}+\frac{m_3\mathbf{r_{23}}}{r^3_{23}}+\frac{m_4\mathbf{r_{24}}}{r^3_{24}}

@@

\ddot{\mathbf{r_3}}=\frac{m_1\mathbf{r_{31}}}{r^3_{31}}+\frac{m_2\mathbf{r_{32}}}{r^3_{32}}+\frac{m_4\mathbf{r_{34}}}{r^3_{34}}

@@

\ddot{\mathbf{r_4}}=\frac{m_1\mathbf{r_{41}}}{r^3_{41}}+\frac{m_2\mathbf{r_{42}}}{r^3_{42}}+\frac{m_3\mathbf{r_{43}}}{r^3_{43}}

@@

A plot of this initial configuration is shown in Fig. \ref{fig:stablebase}.

@@

\includegraphics[width=0.75\textwidth]{./results/stablebase/Orbit.eps}
\caption{Initial stable configuration}
\label{fig:stablebase}

@@

The inner binary is only visible as a point at the centre of the outer binary.

@@

   1 & .002 & .002 & -.5 & .5 & -.001 & .001 & -.5 & .5\\
   2 & .003 & .003 & -.5 & .5 & -.001 & .001 & -.5 & .5\\
   3 & .004 & .004 & -.5 & .5 & -.004 & .004 & -.5 & .5\\
   4 & .004 & .004 & -.55 & .55 & -.004 & .004 & -.5 & .5\\
   5 & .004 & .004 & -.57 & .57 & -.004 & .004 & -.5 & .5\\
   6 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.3 & .3\\
   7 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.4 & .4\\
   8 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.35 & .35\\
   9 & .006 & .006 & -.6 & .6 & -.006 & .006 & -.3 & .3\\
   10 & .01 & .01 & -.7 & .7 & -.01 & .01 & -.2 & .2\\
   11 & .02 & .02 & -.75 & .75 & -.02 & .02 & -.1 & .1\\
   12 & .025 & .025 & -.65 & .65 & -.02 & .02 & -.5 & .5\\
   13 & .05 & .05 & -.9 & .9 & -.05 & .05 & -.1 & .1\\
   14 & .05 & .05 & -.9 & .9 & -.05 & .05 & -.15 & .15\\
   15 & .06 & .06 & -.95 & .95 & -.06 & .06 & -.15 & .15\\
   16 & .07 & .07 & -.1 & .1 & -.07 & .07 & -.15 & .15\\
   17 & .08 & .08 & -.105 & .105 & -.08 & .08 & -.12 & .12\\

@@

\section{Results}
As can be seen from Fig. \ref{fig:stablebase} even a neglible mass inner binary causes the outer binary to deviate from a circular orbit.
Figures Fig. \ref{fig:config1} to \ref{fig:config17} show plots of the double binary using the parameters given in Table \ref{tab:variables}.
As the mass of the inner binary increased the system became increasingly unstable and the effect on the outer binary became more pronounced.
Even the 'stable' configurations listed in Table \ref{tab:variables} hold for only a few orbits.
I was unable to recreate a stable system for a mass $>.08$ for the inner bar.
The plots below reveal that increasing the mass of the inner binary causes the outer binary to separate and
form wider orbits.

\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/002-5-001/Orbit.eps}
\caption{Configuration 1}
\label{fig:config1}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/003-5-001/Orbit.eps}
\caption{Configuration 2}
\label{fig:config2}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-5-004/Orbit.eps}
\caption{Configuration 3}
\label{fig:config3}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-55-004/Orbit.eps}
\caption{Configuration 4}
\label{fig:config4}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-57-004/Orbit.eps}
\caption{Configuration 5}
\label{fig:config5}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-3/Orbit.eps}
\caption{Configuration 6}
\label{fig:config6}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-4/Orbit.eps}
\caption{Configuration 7}
\label{fig:config7}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-35/Orbit.eps}
\caption{Configuration 8}
\label{fig:config8}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/006-6-006-3/Orbit.eps}
\caption{Configuration 9}
\label{fig:config9}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/01-7-01-2/Orbit.eps}
\caption{Configuration 10}
\label{fig:config10}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/02-75-02-1/Orbit.eps}
\caption{Configuration 11}
\label{fig:config11}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/025-65-02/Orbit.eps}
\caption{Configuration 12}
\label{fig:config12}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/05-9-05-1/Orbit.eps}
\caption{Configuration 13}
\label{fig:config13}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/05-9-05-15/Orbit.eps}
\caption{Configuration 14}
\label{fig:config14}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/06-95-06-15/Orbit.eps}
\caption{Configuration 15}
\label{fig:config15}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/07-1-07-15/Orbit.eps}
\caption{Configuration 16}
\label{fig:config16}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/08-105-08-12/Orbit.eps}
\caption{Configuration 17}
\label{fig:config17}
\end{figure}
\end{document}

>>>>

Please see the Compilers subsection for a few notes on compilers - may or may not be relevant to the project.

@@

where $\mathbf{r_i}=\left(x_i,y_i\right)$ and $\mathbf{r_{ij}}=\mathbf{r_j}-\mathbf{r_i}$. So for a four body problem 

@@

\ddot{\mathbf{r_1}}=\frac{m_2\mathbf{r}_{12}}{r^3_{12}}+\frac{m_3\mathbf{r}_{13}}{r^3_{13}}+\frac{m_4\mathbf{r}_{14}}{r^3_{14}}

@@

\ddot{\mathbf{r_2}}=\frac{m_1\mathbf{r}_{21}}{r^3_{21}}+\frac{m_3\mathbf{r}_{23}}{r^3_{23}}+\frac{m_4\mathbf{r}_{24}}{r^3_{24}}

@@

\ddot{\mathbf{r_3}}=\frac{m_1\mathbf{r}_{31}}{r^3_{31}}+\frac{m_2\mathbf{r}_{32}}{r^3_{32}}+\frac{m_4\mathbf{r}_{34}}{r^3_{34}}

@@

\ddot{\mathbf{r_4}}=\frac{m_1\mathbf{r}_{41}}{r^3_{41}}+\frac{m_2\mathbf{r}_{42}}{r^3_{42}}+\frac{m_3\mathbf{r}_{43}}{r^3_{43}}

@@

Plots of the two initial binaries are shown in Fig. \ref{fig:outerbinary} and 
\ref{fig:innerbinary}. The scale of the inner binary is so small that it is difficult
to see in GnuPlot.

@@

\includegraphics[width=0.75\textwidth]{./results/outerbinary/Orbit.eps}
\caption{Outer Binary}
\label{fig:outerbinary}

@@

\includegraphics[width=0.75\textwidth]{./results/innerbinary/Orbit.eps}
\caption{Inner Binary}
\label{fig:innerbinary}

@@

   1 & .001 & .001 & -.5 & .5 & -.001 & .001 & -.5 & .5\\
   2 & .002 & .002 & -.5 & .5 & -.001 & .001 & -.5 & .5\\
   3 & .003 & .003 & -.5 & .5 & -.001 & .001 & -.5 & .5\\
   4 & .004 & .004 & -.5 & .5 & -.004 & .004 & -.5 & .5\\
   5 & .004 & .004 & -.55 & .55 & -.004 & .004 & -.5 & .5\\
   6 & .004 & .004 & -.57 & .57 & -.004 & .004 & -.5 & .5\\
   7 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.3 & .3\\
   8 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.4 & .4\\
   9 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.35 & .35\\
   10 & .006 & .006 & -.6 & .6 & -.006 & .006 & -.3 & .3\\
   11 & .01 & .01 & -.7 & .7 & -.01 & .01 & -.2 & .2\\
   12 & .02 & .02 & -.75 & .75 & -.02 & .02 & -.1 & .1\\
   13 & .025 & .025 & -.65 & .65 & -.02 & .02 & -.5 & .5\\
   14 & .05 & .05 & -.9 & .9 & -.05 & .05 & -.1 & .1\\
   15 & .05 & .05 & -.9 & .9 & -.05 & .05 & -.15 & .15\\
   16 & .06 & .06 & -.95 & .95 & -.06 & .06 & -.15 & .15\\
   17 & .07 & .07 & -.1 & .1 & -.07 & .07 & -.15 & .15\\
   18 & .08 & .08 & -.105 & .105 & -.08 & .08 & -.12 & .12\\

@@

\subsection{Compilers}
As I had converted OdeInt into Fortran 90 I wanted to make sure I hadn't introduced any errors. I therefore compared the output of the Numerical Recipes
version of OdeInt with mine. I was compiling with GFortran and the Fortran 77 version was compiled with g77. They were different! So naturally concerned 
I then compiled the Fortran 77 version with GFortran. The results were identical to mine, which satisified me that I hadn't introduced any errors.

\section{Results}
As can be seen from Fig. \ref{fig:config1} even a neglible mass inner binary causes the outer binary to begin to separate and become less tightly bound.
Figures Fig. \ref{fig:config1} to \ref{fig:config18} show plots of the double binary using the parameters given in Table \ref{tab:variables}.
In many cases the inner binary is only visible as a point at the centre of the outer binary.
As the mass of the inner binary increased the system became increasingly unstable and the effect on the outer binary became more pronounced.
Even the 'stable' configurations listed in Table \ref{tab:variables} hold for only a few orbits.
I was unable to recreate a stable system for a mass $>.08$ for the inner bar.
The plots below reveal that increasing the mass of the inner binary causes the outer binary to separate and
form wider orbits.
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/stablebase/Orbit.eps}
\caption{Configuration 1}
\label{fig:config1}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/002-5-001/Orbit.eps}
\caption{Configuration 2}
\label{fig:config2}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/003-5-001/Orbit.eps}
\caption{Configuration 3}
\label{fig:config3}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-5-004/Orbit.eps}
\caption{Configuration 4}
\label{fig:config4}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-55-004/Orbit.eps}
\caption{Configuration 5}
\label{fig:config5}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-57-004/Orbit.eps}
\caption{Configuration 6}
\label{fig:config6}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-3/Orbit.eps}
\caption{Configuration 7}
\label{fig:config7}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-4/Orbit.eps}
\caption{Configuration 8}
\label{fig:config8}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-35/Orbit.eps}
\caption{Configuration 9}
\label{fig:config9}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/006-6-006-3/Orbit.eps}
\caption{Configuration 10}
\label{fig:config10}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/01-7-01-2/Orbit.eps}
\caption{Configuration 11}
\label{fig:config11}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/02-75-02-1/Orbit.eps}
\caption{Configuration 12}
\label{fig:config12}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/025-65-02/Orbit.eps}
\caption{Configuration 13}
\label{fig:config13}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/05-9-05-1/Orbit.eps}
\caption{Configuration 14}
\label{fig:config14}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/05-9-05-15/Orbit.eps}
\caption{Configuration 15}
\label{fig:config15}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/06-95-06-15/Orbit.eps}
\caption{Configuration 16}
\label{fig:config16}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/07-1-07-15/Orbit.eps}
\caption{Configuration 17}
\label{fig:config17}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/08-105-08-12/Orbit.eps}
\caption{Configuration 18}
\label{fig:config18}
\end{figure}
\end{document}
